---
title: "prop_cum_method"
output: html_notebook
---


Identify samples which have a significant change in average exon expression based on the  cumulative counts function which is run for all ten test genes across corresponding cancer types (Tables 1 and 2, Supplementary Table 8).

```{r}
#For 3' genes on the negative strand
get_all_sig_z_prop_cum <- function(gene, project_id) {
  df <- raw_exon_counts_neg(gene)
  proj_df <- subset_project_neg(gene, project_id, df)
  cumsum_df <- cumsum_calc_neg(gene, proj_df)
  diff_df <- diff_calc(cumsum_df)
  diff_auc_df <- norm_auc_neg(gene, project_id, diff_df)
  all_z <- z_calc(diff_auc_df)
  all_z
}

#For 3' genes on the positive strand
get_all_sig_z_prop_cum_pos <- function(gene, project_id) {
  df <- raw_exon_counts(gene)
  proj_df <- subset_project(gene, project_id, df)
  cumsum_df <- cumsum_calc(gene, proj_df)
  diff_df <- diff_calc(cumsum_df)
  diff_auc_df <- norm_auc(gene, project_id, diff_df)
  all_z <- z_calc(diff_auc_df)
  all_z
}
```

All ten test genes by corresponding cancer type:
```{r}
alk_blca_prop_cum_all <- get_all_sig_z_prop_cum("ALK", "TCGA-BLCA")
alk_kirp_prop_cum_all <- get_all_sig_z_prop_cum("ALK", "TCGA-KIRP")
alk_luad_prop_cum_all <- get_all_sig_z_prop_cum("ALK", "TCGA-LUAD")
alk_read_prop_cum_all <- get_all_sig_z_prop_cum("ALK", "TCGA-READ")
alk_sarc_prop_cum_all <- get_all_sig_z_prop_cum("ALK", "TCGA-SARC")
alk_skcm_prop_cum_all <- get_all_sig_z_prop_cum("ALK", "TCGA-SKCM")
alk_thca_prop_cum_all <- get_all_sig_z_prop_cum("ALK", "TCGA-THCA")

erg_cesc_prop_cum_all <- get_all_sig_z_prop_cum("ERG", "TCGA-CESC")
erg_lgg_prop_cum_all <- get_all_sig_z_prop_cum("ERG", "TCGA-LGG")
erg_paad_prop_cum_all <- get_all_sig_z_prop_cum("ERG", "TCGA-PAAD")
erg_pcpg_prop_cum_all <- get_all_sig_z_prop_cum("ERG", "TCGA-PCPG")
erg_prad_prop_cum_all <- get_all_sig_z_prop_cum("ERG", "TCGA-PRAD")

etv1_lgg_prop_cum_all <- get_all_sig_z_prop_cum("ETV1", "TCGA-LGG")
etv1_prad_prop_cum_all <- get_all_sig_z_prop_cum("ETV1", "TCGA-PRAD")

maml3_brca_prop_cum_all <- get_all_sig_z_prop_cum("MAML3", "TCGA-BRCA")
maml3_lihc_prop_cum_all <- get_all_sig_z_prop_cum("MAML3", "TCGA-LIHC")
maml3_pcpg_prop_cum_all <- get_all_sig_z_prop_cum("MAML3", "TCGA-PCPG")

ntrk3_brca_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-BRCA")
ntrk3_cesc_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-CESC")
ntrk3_coad_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-COAD")
ntrk3_hnsc_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-HNSC")
ntrk3_lgg_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-LGG")
ntrk3_paad_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-PAAD")
ntrk3_skcm_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-SKCM")
ntrk3_thca_prop_cum_all <- get_all_sig_z_prop_cum("NTRK3", "TCGA-THCA")

tfe3_kirc_prop_cum_all <- get_all_sig_z_prop_cum("TFE3", "TCGA-KIRC")
tfe3_kirp_prop_cum_all <- get_all_sig_z_prop_cum("TFE3", "TCGA-KIRP")
tfe3_ucec_prop_cum_all <- get_all_sig_z_prop_cum("TFE3", "TCGA-UCEC")

arhgap26_cesc_prop_cum_all <- get_all_sig_z_prop_cum_pos("ARHGAP26", "TCGA-CESC")
arhgap26_hnsc_prop_cum_all <- get_all_sig_z_prop_cum_pos("ARHGAP26", "TCGA-HNSC")
arhgap26_luad_prop_cum_all <- get_all_sig_z_prop_cum_pos("ARHGAP26", "TCGA-LUAD")
arhgap26_lusc_prop_cum_all <- get_all_sig_z_prop_cum_pos("ARHGAP26", "TCGA-LUSC")
arhgap26_sarc_prop_cum_all <- get_all_sig_z_prop_cum_pos("ARHGAP26", "TCGA-SARC")
arhgap26_stad_prop_cum_all <- get_all_sig_z_prop_cum_pos("ARHGAP26", "TCGA-STAD")

rara_brca_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-BRCA")
rara_cesc_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-CESC")
rara_gbm_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-GBM")
rara_laml_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-LAML")
rara_ov_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-OV")
rara_pcpg_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-PCPG")
rara_read_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-READ")
rara_stad_prop_cum_all <- get_all_sig_z_prop_cum_pos("RARA", "TCGA-STAD")

ret_brca_prop_cum_all <- get_all_sig_z_prop_cum_pos("RET", "TCGA-BRCA")
ret_coad_prop_cum_all <- get_all_sig_z_prop_cum_pos("RET", "TCGA-COAD")
ret_laml_prop_cum_all <- get_all_sig_z_prop_cum_pos("RET", "TCGA-LAML")
ret_luad_prop_cum_all <- get_all_sig_z_prop_cum_pos("RET", "TCGA-LUAD")
ret_ov_prop_cum_all <- get_all_sig_z_prop_cum_pos("RET", "TCGA-OV")
ret_thca_prop_cum_all <- get_all_sig_z_prop_cum_pos("RET", "TCGA-THCA")

tacc3_blca_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-BLCA")
tacc3_brca_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-BRCA")
tacc3_cesc_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-CESC")
tacc3_esca_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-ESCA")
tacc3_gbm_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-GBM")
tacc3_hnsc_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-HNSC")
tacc3_kirp_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-KIRP")
tacc3_laml_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-LAML")
tacc3_lgg_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-LGG")
tacc3_lihc_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-LIHC")
tacc3_luad_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-LUAD")
tacc3_lusc_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-LUSC")
tacc3_stad_prop_cum_all <- get_all_sig_z_prop_cum_pos("TACC3", "TCGA-STAD")
```
Merge each 3' gene across cancer types into one dataframe per cancer type. Saved as Rdata file prop_cum_all.Rdata
```{r}
all_alk_prop_cum_all <- cbind(alk_blca_prop_cum_all,
alk_kirp_prop_cum_all,
alk_luad_prop_cum_all,
alk_read_prop_cum_all,
alk_sarc_prop_cum_all,
alk_skcm_prop_cum_all,
alk_thca_prop_cum_all)

all_arhgap26_prop_cum_all <- cbind(arhgap26_cesc_prop_cum_all,
arhgap26_hnsc_prop_cum_all,
arhgap26_luad_prop_cum_all,
arhgap26_lusc_prop_cum_all,
arhgap26_sarc_prop_cum_all,
arhgap26_stad_prop_cum_all)

all_erg_prop_cum_all <- cbind(erg_cesc_prop_cum_all,
erg_lgg_prop_cum_all,
erg_paad_prop_cum_all,
erg_pcpg_prop_cum_all,
erg_prad_prop_cum_all)

all_etv1_prop_cum_all <- cbind(etv1_lgg_prop_cum_all,
etv1_prad_prop_cum_all)

all_maml3_prop_cum_all <- cbind(maml3_brca_prop_cum_all,
maml3_lihc_prop_cum_all,
maml3_pcpg_prop_cum_all)

all_ntrk3_prop_cum_all <- cbind(ntrk3_brca_prop_cum_all,
ntrk3_cesc_prop_cum_all,
ntrk3_coad_prop_cum_all,
ntrk3_hnsc_prop_cum_all,
ntrk3_lgg_prop_cum_all,
ntrk3_paad_prop_cum_all,
ntrk3_skcm_prop_cum_all,
ntrk3_thca_prop_cum_all)

all_rara_prop_cum_all <- cbind(rara_brca_prop_cum_all,
rara_cesc_prop_cum_all,
rara_gbm_prop_cum_all,
rara_laml_prop_cum_all,
rara_ov_prop_cum_all,
rara_pcpg_prop_cum_all,
rara_read_prop_cum_all,
rara_stad_prop_cum_all)

all_ret_prop_cum_all <- cbind(ret_brca_prop_cum_all,
ret_coad_prop_cum_all,
ret_laml_prop_cum_all,
ret_luad_prop_cum_all,
ret_ov_prop_cum_all,
ret_thca_prop_cum_all)

all_tacc3_prop_cum_all <- cbind(tacc3_blca_prop_cum_all,
tacc3_brca_prop_cum_all,
tacc3_cesc_prop_cum_all,
tacc3_esca_prop_cum_all,
tacc3_gbm_prop_cum_all,
tacc3_hnsc_prop_cum_all,
tacc3_kirp_prop_cum_all,
tacc3_laml_prop_cum_all,
tacc3_lgg_prop_cum_all,
tacc3_lihc_prop_cum_all,
tacc3_luad_prop_cum_all,
tacc3_lusc_prop_cum_all,
tacc3_stad_prop_cum_all)

all_tfe3_prop_cum_all <- cbind(tfe3_kirc_prop_cum_all,
tfe3_kirp_prop_cum_all,
tfe3_ucec_prop_cum_all)

```

Get true positives for each gene at recount3 exon as per Table 2.

```{r}
get_tp <- function(gene, df) {
  ids <- get_known_ids_new(gene) #Generate rail_ids known to contain fusion genes
  subset(df, rownames(df) %in% ids) #Subset to known fusion gene samples
}

#df input Z scores
diff_sub_all_ids(df, n) {
  new_df <- as.data.frame(df[n,] > 3) #n is the most common fusion breakpoint, see Table 2 
  all_ids <- as.data.frame(apply(
    new_df, 1, 
    function(u) paste( names(which(u))) 
  ))
  rownames(all_ids) <- all_ids[,1]
  colnames(all_ids) <- "Significant"
  all_ids
}
```

And false positives from the get_tp output, using base R setdiff(), e.g. 
```{r}
setdiff(all_tfe3_prop_cum_all, get_tp(all_tfe3_prop_cum_all)) #Fetches significant samples that do not contain fusion genes
```

The full list of true positives is available from file truepos_prop_cum.csv
The full list of false positives is avaiable from file falsepos_prop_cum.csv

