---
title: "R Notebook"
output: html_notebook
---

TCGA samples for the ten test genes across their corresponding cancer types (as per Tables 1 and 2, and Supplementary Table 8) determined to have a significant change in individual exon expression at the most common fusion breakpoint (Table 2). 

The samples were calculated using function, which generated all true and false positive values.
```{r}
#For 3' genes on the negative strand
function(gene, project_id) {
  df <- raw_exon_counts_neg(gene)
  proj_df <- subset_project_neg(gene, project_id, df)
  kbp_df <- norm_length_neg(gene, proj_df)
  diff_df <- diff_calc(kbp_df)
  diff_auc_df <- norm_auc_neg(gene, project_id, diff_df)
  all_z <- z_calc(diff_auc_df)
  all_z
}

#For 3' genes on the positive strand
get_all_sig_z_diff_pos <- function(gene, project_id) {
  df <- raw_exon_counts(gene)
  proj_df <- subset_project(gene, project_id, df)
  kbp_df <- norm_length(gene, proj_df)
  diff_df <- diff_calc(kbp_df)
  diff_auc_df <- norm_auc(gene, project_id, diff_df)
  all_z <- z_calc(diff_auc_df)
  all_z
}
```

All ten test genes by corresponding cancer type:
```{r}
alk_blca_diff_all <- get_all_sig_z_diff("ALK", "TCGA-BLCA")
alk_kirp_diff_all <- get_all_sig_z_diff("ALK", "TCGA-KIRP")
alk_luad_diff_all <- get_all_sig_z_diff("ALK", "TCGA-LUAD")
alk_read_diff_all <- get_all_sig_z_diff("ALK", "TCGA-READ")
alk_sarc_diff_all <- get_all_sig_z_diff("ALK", "TCGA-SARC")
alk_skcm_diff_all <- get_all_sig_z_diff("ALK", "TCGA-SKCM")
alk_thca_diff_all <- get_all_sig_z_diff("ALK", "TCGA-THCA")

erg_cesc_diff_all <- get_all_sig_z_diff("ERG", "TCGA-CESC")
erg_lgg_diff_all <- get_all_sig_z_diff("ERG", "TCGA-LGG")
erg_paad_diff_all <- get_all_sig_z_diff("ERG", "TCGA-PAAD")
erg_pcpg_diff_all <- get_all_sig_z_diff("ERG", "TCGA-PCPG")
erg_prad_diff_all <- get_all_sig_z_diff("ERG", "TCGA-PRAD")

etv1_lgg_diff_all <- get_all_sig_z_diff("ETV1", "TCGA-LGG")
etv1_prad_diff_all <- get_all_sig_z_diff("ETV1", "TCGA-PRAD")

maml3_brca_diff_all <- get_all_sig_z_diff("MAML3", "TCGA-BRCA")
maml3_lihc_diff_all <- get_all_sig_z_diff("MAML3", "TCGA-LIHC")
maml3_pcpg_diff_all <- get_all_sig_z_diff("MAML3", "TCGA-PCPG")

ntrk3_brca_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-BRCA")
ntrk3_cesc_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-CESC")
ntrk3_coad_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-COAD")
ntrk3_hnsc_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-HNSC")
ntrk3_lgg_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-LGG")
ntrk3_paad_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-PAAD")
ntrk3_skcm_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-SKCM")
ntrk3_thca_diff_all <- get_all_sig_z_diff("NTRK3", "TCGA-THCA")

tfe3_kirc_diff_all <- get_all_sig_z_diff("TFE3", "TCGA-KIRC")
tfe3_kirp_diff_all <- get_all_sig_z_diff("TFE3", "TCGA-KIRP")
tfe3_ucec_diff_all <- get_all_sig_z_diff("TFE3", "TCGA-UCEC")

arhgap26_cesc_diff_all <- get_all_sig_z_diff_pos("ARHGAP26", "TCGA-CESC")
arhgap26_hnsc_diff_all <- get_all_sig_z_diff_pos("ARHGAP26", "TCGA-HNSC")
arhgap26_luad_diff_all <- get_all_sig_z_diff_pos("ARHGAP26", "TCGA-LUAD")
arhgap26_lusc_diff_all <- get_all_sig_z_diff_pos("ARHGAP26", "TCGA-LUSC")
arhgap26_sarc_diff_all <- get_all_sig_z_diff_pos("ARHGAP26", "TCGA-SARC")
arhgap26_stad_diff_all <- get_all_sig_z_diff_pos("ARHGAP26", "TCGA-STAD")

rara_brca_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-BRCA")
rara_cesc_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-CESC")
rara_gbm_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-GBM")
rara_laml_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-LAML")
rara_ov_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-OV")
rara_pcpg_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-PCPG")
rara_read_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-READ")
rara_stad_diff_all <- get_all_sig_z_diff_pos("RARA", "TCGA-STAD")

ret_brca_diff_all <- get_all_sig_z_diff_pos("RET", "TCGA-BRCA")
ret_coad_diff_all <- get_all_sig_z_diff_pos("RET", "TCGA-COAD")
ret_laml_diff_all <- get_all_sig_z_diff_pos("RET", "TCGA-LAML")
ret_luad_diff_all <- get_all_sig_z_diff_pos("RET", "TCGA-LUAD")
ret_ov_diff_all <- get_all_sig_z_diff_pos("RET", "TCGA-OV")
ret_thca_diff_all <- get_all_sig_z_diff_pos("RET", "TCGA-THCA")

tacc3_blca_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-BLCA")
tacc3_brca_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-BRCA")
tacc3_cesc_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-CESC")
tacc3_esca_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-ESCA")
tacc3_gbm_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-GBM")
tacc3_hnsc_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-HNSC")
tacc3_kirp_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-KIRP")
tacc3_laml_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-LAML")
tacc3_lgg_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-LGG")
tacc3_lihc_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-LIHC")
tacc3_luad_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-LUAD")
tacc3_lusc_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-LUSC")
tacc3_stad_diff_all <- get_all_sig_z_diff_pos("TACC3", "TCGA-STAD")
```

Merge each 3' gene across cancer types into one dataframe per cancer type. Saved as Rdata file all_diff_objects.Rdata
```{r}
all_alk_diff_all <- cbind(alk_blca_diff_all,
alk_kirp_diff_all,
alk_luad_diff_all,
alk_read_diff_all,
alk_sarc_diff_all,
alk_skcm_diff_all,
alk_thca_diff_all)

all_arhgap26_diff_all <- cbind(arhgap26_cesc_diff_all,
arhgap26_hnsc_diff_all,
arhgap26_luad_diff_all,
arhgap26_lusc_diff_all,
arhgap26_sarc_diff_all,
arhgap26_stad_diff_all)

all_erg_diff_all <- cbind(erg_cesc_diff_all,
erg_lgg_diff_all,
erg_paad_diff_all,
erg_pcpg_diff_all,
erg_prad_diff_all)

all_etv1_diff_all <- cbind(etv1_lgg_diff_all,
etv1_prad_diff_all)

all_maml3_diff_all <- cbind(maml3_brca_diff_all,
maml3_lihc_diff_all,
maml3_pcpg_diff_all)

all_ntrk3_diff_all <- cbind(ntrk3_brca_diff_all,
ntrk3_cesc_diff_all,
ntrk3_coad_diff_all,
ntrk3_hnsc_diff_all,
ntrk3_lgg_diff_all,
ntrk3_paad_diff_all,
ntrk3_skcm_diff_all,
ntrk3_thca_diff_all)

all_rara_diff_all <- cbind(rara_brca_diff_all,
rara_cesc_diff_all,
rara_gbm_diff_all,
rara_laml_diff_all,
rara_ov_diff_all,
rara_pcpg_diff_all,
rara_read_diff_all,
rara_stad_diff_all)

all_ret_diff_all <- cbind(ret_brca_diff_all,
ret_coad_diff_all,
ret_laml_diff_all,
ret_luad_diff_all,
ret_ov_diff_all,
ret_thca_diff_all)

all_tacc3_diff_all <- cbind(tacc3_blca_diff_all,
tacc3_brca_diff_all,
tacc3_cesc_diff_all,
tacc3_esca_diff_all,
tacc3_gbm_diff_all,
tacc3_hnsc_diff_all,
tacc3_kirp_diff_all,
tacc3_laml_diff_all,
tacc3_lgg_diff_all,
tacc3_lihc_diff_all,
tacc3_luad_diff_all,
tacc3_lusc_diff_all,
tacc3_stad_diff_all)

all_tfe3_diff_all <- cbind(tfe3_kirc_diff_all,
tfe3_kirp_diff_all,
tfe3_ucec_diff_all)
```

Get true  and false positives for each gene at recount3 exon as per Table 2. See average_exon_values_thesis.Rmd for process.

The full list of true positives is available from file truepos_diff.csv
The full list of false positives is avaiable from file falsepos_diff.csv

