---
title: "Workflow"
output: html_notebook
---
```{r}
library(recount3)
library(snapcount)
library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyverse)
library(EnvStats) 
library(gsubfn)
```


##Gene overexpression workflow using TACC3 as an example

#Generating count dataframe
```{r}
#gene query functions
neg_gene_query <- function(gene) {
  sb <- QueryBuilder(compilation = "tcga", regions = gene)
  sb <- set_row_filters(sb, strand == "-")
  sb <- query_gene(sb)
  sb
}

pos_gene_query <- function(gene) {
  sb <- QueryBuilder(compilation = "tcga", regions = gene)
  sb <- set_row_filters(sb, strand == "+")
  sb <- query_gene(sb)
  sb
}

gtex_neg_gene_query <- function(gene) {
  sb <- QueryBuilder(compilation = "gtex", regions = gene)
  sb <- set_row_filters(sb, strand == "-")
  sb <- query_gene(sb)
  sb
}

gtex_pos_gene_query <- function(gene) {
  sb <- QueryBuilder(compilation = "gtex", regions = gene)
  sb <- set_row_filters(sb, strand == "+")
  sb <- query_gene(sb)
  sb
}


#functions to generate count dataframes
tcga_ids <- read.csv('tcga.samples.csv') 

gene_counts <- function(gene, df, gene.ID) {
  gene_counts <- as.data.frame(as.matrix(assays(gene)$counts)[1,])
  colnames(gene_counts) <- "Counts"
  gene_counts$rail_id <- colData(gene)[, "rail_id"]
  gene_counts <- gene_counts %>% transform(rail_id = sub("^", "rail_", gene_counts$rail_id))
  gene_counts$Cancer <- colData(gene)[,"gdc_cases.project.name"]
  gene_counts$Tissue <- colData(gene)[, "gdc_cases.project.primary_site"]
  gene_counts$Tissue <- recode(gene_counts$Tissue,
                               "Bone Marrow" = "Blood")
  gene_counts$Project_ID <- colData(gene)[, "gdc_cases.project.project_id"]
  gene_counts <- gene_counts %>% transform(Project_ID = gsub("TCGA-", "", gene_counts$Project_ID))
  gene_counts$ID <- colData(gene)[, "gdc_cases.samples.portions.analytes.aliquots.submitter_id"]
  gene_counts$auc <- colData(gene)[, "auc"]
  gene_counts$mapped <- colData(gene)[, "mapped_read_count"]
  gene_counts <- gene_counts %>% mutate(auc_norm = (Counts/auc)*1000000, mapped_norm = (Counts/mapped)*1000000, log2_Counts = log2(Counts), log2_auc_norm = log2(auc_norm), log2_map_norm = log2(mapped_norm))
  gene_counts$has_cancer <- ifelse(colData(gene)[, "gdc_cases.samples.sample_type"] == "Solid Tissue Normal", "Normal", "Cancer")
  projects <- c("ACC", "PCPG", "GBM", "LGG", "READ", "COAD", "KIRP", "KIRC", "KICH", "LUSC", "LUAD", "UCEC", "UCS")
  projects <- data.frame(projects)
  cancer <- gene_counts %>% filter(has_cancer == "Cancer") %>% mutate(cancer_status = ifelse(Project_ID %in% projects[,1], .$Project_ID, "Cancer"))
  normal <- gene_counts %>% filter(has_cancer == "Normal") 
  normal$cancer_status <- "Normal"
  gene_counts2 <- rbind(cancer, normal)
  gene_counts2 <- gene_counts2[order(gene_counts2$log2_auc_norm, decreasing = TRUE),] 
  gene_counts2 %>%
  mutate(fusion = ifelse(gene_counts2$ID %in% df[, gene.ID], "Fusion", "No Fusion")) 
}

gtex_counts <- function(gene) {
  gtex_counts <- as.data.frame(as.matrix(assays(gene)$counts)[1,])
  colnames(gtex_counts) <- "Counts"
  gtex_counts$rail_id <- colData(gene)[, "rail_id"]
  gtex_counts <- gtex_counts %>% transform(rail_id = sub("^", "rail_", gtex_counts$rail_id))
  gtex_counts$Cancer <- "GTEx"
  gtex_counts$Tissue <- colData(gene)[,"Histological_Type"]
  gtex_counts$Tissue <- recode(gtex_counts$Tissue, 
                              "Cervix Uteri" = "Cervix")
  gtex_counts$Project_ID <- "GTEx"
  gtex_counts$ID <- "GTEx"
  gtex_counts$auc <- colData(gene)[, "auc"]
  gtex_counts$mapped <- colData(gene)[, "mapped_read_count"]
  gtex_counts <- gtex_counts %>% mutate(auc_norm = (Counts/auc)*1000000, mapped_norm = (Counts/mapped)*1000000, log2_Counts = log2(Counts), log2_auc_norm = log2(auc_norm), log2_map_norm = log2(mapped_norm))
  gtex_counts$has_cancer <- "GTEx"
  gtex_counts$cancer_status <- "GTEx"
  gtex_counts$fusion <- "GTEx"
  gtex_counts <- gtex_counts[order(gtex_counts$log2_auc_norm, decreasing = TRUE),]
  gtex_counts
}

#function to generate combined dataframe
combined_counts <- function(gene_counts, gtex_counts) {
  combined_counts <- rbind(gene_counts, gtex_counts)
  combined_counts <- combined_counts[, -c(7:8)]
  combined_counts <- combined_counts %>% filter(Tissue != "Bone Marrow", 
                                                Tissue != "Vagina", 
                                                Tissue != "Spleen", 
                                                Tissue != "Fallopian Tube", 
                                                Tissue != "NA",
                                                Tissue != "Adipose Tissue",
                                                Tissue != "Muscle",
                                                Tissue != "Blood Vessel",
                                                Tissue != "Nerve",
                                                Tissue != "Colon",
                                                Tissue != "Small Intestine",
                                                Tissue != "Pituitary",
                                                Tissue != "Heart",
                                                Tissue != "Salivary Gland")
  combined_counts <- combined_counts %>%
    group_by(Tissue, cancer_status) %>%
    mutate(min = min(log2_auc_norm),
            quant25 = quantile(log2_auc_norm, probs = 0.25), 
            quant50 = quantile(log2_auc_norm, probs = 0.5),
            quant75 = quantile(log2_auc_norm, probs = 0.75),
            quant90 = quantile(log2_auc_norm, probs = 0.9),
            quant95 = quantile(log2_auc_norm, probs = 0.95),
            max = max(log2_auc_norm),
            outlier = (quant75 + 1.5*(quant75 - quant25)))
  combined_counts <- combined_counts %>%
    group_by(Tissue, cancer_status) %>%
    mutate(is_outlier = ifelse(log2_auc_norm >= outlier, T, F),
           above_90 = ifelse(log2_auc_norm >= quant90, T, F),
           above_95 = ifelse(log2_auc_norm >= quant95, T, F))
  combined_counts %>% mutate(is_expressed = ifelse(quant25 <= 1, "Not normally expressed", "Normally expressed")) 
}
```

```{r}
tacc3_gene <- pos_gene_query("tacc3")
gtex_tacc3_gene <- gtex_pos_gene_query("tacc3")
tacc3_counts <- gene_counts(tacc3_gene, tcga_ids, "TACC3.ID")
gtex_tacc3_counts <- gtex_counts(gtex_tacc3_gene)
tacc3_combined <- combined_counts(tacc3_counts, gtex_tacc3_counts)
```

#identifying overexpressed samples
```{r}
gene_potential_fusions <- tacc3_combined %>% 
  filter(has_cancer == "Cancer", above_95 == TRUE) %>%
  ungroup() %>% 
  group_by(Tissue, Project_ID) %>%
  select(rail_id, Tissue, Project_ID, Cancer, ID, auc_norm)
gene_potential_fusions
write.csv(gene_potential_fusions, 'gene_overexp_tacc3_potential_fusions.csv')
```


#comparing TCGA cancer and GTEx data
one sided Mann-Whitney-Wilcoxon test (TCGA > GTEx)
```{r}
compare_tcga_gtex_sided <- function(gene) {
  counts <- gene
  counts <- counts %>% 
  filter(has_cancer != "Normal") %>% 
  filter(Tissue != "Bile Duct", Tissue != "Colorectal", Tissue != "Eye", Tissue != "Head and Neck", Tissue != "Lymph Nodes", Tissue != "Pleura", Tissue != "Soft Tissue", Tissue != "Thymus") %>%
  group_by(has_cancer) %>%
  split(f = list(as.character(.$Tissue))) %>%
  map(~ wilcox.test(log2_auc_norm ~ has_cancer, data = .x, alternative = "greater")$p.value) %>%
  unlist(., .$Tissue) %>%
  data.frame(.)
  counts
}
compare_tcga_gtex_sided(tacc3_combined)
```
-TCGA tissues that are significant (p < 0.05)
```{r}
compare_tcga_gtex_sided(tacc3_combined) %>% filter(. < 0.05)
```

boxplot visualisation with samples above 95th percentile overlaid in grey
```{r}
tacc3_combined %>%
  filter(has_cancer == "Cancer") %>%
  ggplot(aes(x = Project_ID, y = log2_auc_norm, fill = Project_ID)) +
  geom_boxplot() +
  geom_jitter(data = . %>% filter(above_95 == TRUE), colour = "darkgrey") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))
```
split by tissue and data source
```{r}
tacc3_combined %>%
  ggplot(aes(x = has_cancer, y = log2_auc_norm, fill = has_cancer)) +
  geom_boxplot() +
  geom_jitter(data = . %>% filter(has_cancer == "Cancer", above_95 == TRUE), colour = "darkgrey") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)) +
  facet_wrap(~ Tissue)
```





